%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Advanced Topics in the rmongodb Package}

Advanced Topics in the rmongodb Package
========================================

MongoDB (www.mongodb.org) is a scalable, high-performance, document-oriented NoSQL database. The **rmongodb** package provides an interface from the statistical software R (www.r-project.org) to MongoDB and back using the mongodb-C library. 


This vignette will provide demos for advanced topics in the **rmongodb** package.



# Connect R to MongoDB
First of all we have to load the library and to connect to a MongoDB. In this case we connect to our local MongoDB installation.
```{r connect}
library(rmongodb)
mongo <- mongo.create()
mongo.is.connected(mongo)
```




# Insert Big Data
We will use the "Zip Code Data Set" from MongoDB in the following examples. The data set is included in the **rmongodb** package and can be loaded using the command *data(zips)*. The 'zips' data set holds the MongoDB example data set called "Zip Code Data Set" (http://docs.mongodb.org/manual/tutorial/aggregation-zip-code-data-set/). This data set is available as JSON and hold zip code data from the US.
```{r insertZips}
# load example data set from rmongodb
data(zips)
head(zips)
zips[1,]$loc
  
if(mongo.is.connected(mongo) == TRUE){
  res <- list(length(dim(zips)[1]))
  for(i in 1:dim(zips)[1]){
    tmp <- zips[i,] 
    res[[i]] <- mongo.bson.from.list(tmp)
  }
  mongo.insert.batch(mongo, "rmongodb.zips", res )
}
```
Lets do a small check to test if all inserted data is available in MongoDB.
```{r insertCheck}
dim(zips)
if(mongo.is.connected(mongo) == TRUE){
  nr <- mongo.count(mongo, "rmongodb.zips")
  print( nr )
  res <- mongo.find.all(mongo, "rmongodb.zips")
  head(res)
}
```



# MongoDB Aggregation Framework
MongoDB Aggregation Framework is one of the top MongoDB features. Aggregations operations process data records and return computed results. Aggregation operations group values from multiple documents together, and can perform a variety of operations on the grouped data to return a single result. 

So far, JSON to BSON conversion is not yet working for aggregation requests in **rmongodb**. Therefore, we have to create these difficult BSON objects.
```{r aggregationFrameworkBSON, results='hide'}
buf <- mongo.bson.buffer.create()
mongo.bson.buffer.start.object(buf, "$group")
mongo.bson.buffer.append(buf, "_id", "$state")
mongo.bson.buffer.start.object(buf, "totalPop")
mongo.bson.buffer.append(buf, "$sum", "$pop")
mongo.bson.buffer.finish.object(buf)
mongo.bson.buffer.finish.object(buf)
bson <- mongo.bson.from.buffer(buf)

bufall <- mongo.bson.buffer.create()
mongo.bson.buffer.append(bufall, "aggregate", "zips")
mongo.bson.buffer.start.array(bufall, "pipeline")
mongo.bson.buffer.append(bufall, "0", bson)
mongo.bson.buffer.finish.object(bufall)
cmd <- mongo.bson.from.buffer(bufall)
```
```{r aggregationFrameworkrun}
cmd
if(mongo.is.connected(mongo) == TRUE){
  res <- mongo.command(mongo, "rmongodb", cmd)
  res
}
```






# GridFS with rmongodb
GridFS is a specification for storing and retrieving files that exceed the BSON-document size limit of 16MB. There are several commands in **rmongodb** available to work with gridFS.
```{r gridfs}
if(mongo.is.connected(mongo) == TRUE){
  mgrids <- mongo.gridfs.create(mongo, "rmongodb", prefix = "fs")
  mongo.gridfs.store.file(mgrids, "faust.txt", "Faust")
  gf <- mongo.gridfs.find(mgrids, "Faust")
  mongo.gridfile.get.length(gf)
  mongo.gridfile.get.chunk.count(gf)
}
```



# Drop / Remove collections and databases and Close Connection to MongoDB
At the end of all your analyses it is a good idea to destroy the connection to your MongoDB and to clean-up collections.
```{r drop}
if(mongo.is.connected(mongo) == TRUE){
  mongo.drop(mongo, "rmongodb.zips")
  mongo.drop.database(mongo, "rmongodb")
  
  # close connection
  mongo.destroy(mongo)
}
```


# Feedback and Issues
Please do not hesitate to contact us if there are any issues using **rmongodb**. Issues or pull requests can be submitted via github: https://github.com/mongosoup/rmongodb


