%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction to the rmongodb Package}

Introduction to the rmongodb Package
========================================

MongoDB (www.mongodb.org) is a scalable, high-performance, document-oriented NoSQL database. The **rmongodb** package provides an interface from the statistical software R (www.r-project.org) to MongoDB and back using the mongodb-C library. 

# Installing and Loading rmongodb Package
There is a stable CRAN version of **rmongodb** available:
```{r installRmongodb, eval=FALSE}
install.packages("rmongodb")
```

You can also install the latest development version from the GitHub repository:
```{r installDEV, eval=FALSE}
library(devtools)
install_github("rmongodb", "mongosoup")
```
The installation should be very simple and straightforward. No local MongoDB installation is required. Only if you install from source the RUnit tests will need a local MongoDB installation.

Afterwards you can load the **rmongodb** package as any other package:
```{r loadRmongodb}
library(rmongodb)
```




# First Steps in rmongodb Package

## Connect to MongoDB
```{r connect2Mongo}
mongo <- mongo.create()
mongo
mongo.is.connected(mongo)
```

```{r importZIPdata, echo=FALSE, warning=FALSE, results='hide'}
if(mongo.is.connected(mongo) == TRUE){
  # load some data
  library(jsonlite)
  data(zips)
  res <- list(length(dim(zips)[1]))
  for(i in 1:dim(zips)[1]){
    tmp <- zips[i,] 
    res[[i]] <- mongo.bson.from.list(tmp)
  }
  mongo.insert.batch(mongo, "rmongodb.zips", res )
}
```

## Get Databases and Collections
```{r getDBsColls}
if(mongo.is.connected(mongo) == TRUE){
  mongo.get.databases(mongo)
  
  db <- "rmongodb"
  coll <- "rmongodb.zips"
  mongo.get.database.collections(mongo, db)
}
```

## Get Size of Collection and Values for a Key
```{r countDistinct}
if(mongo.is.connected(mongo) == TRUE){
  mongo.count(mongo, coll)

  res <- mongo.distinct(mongo, coll, "city")
  head(res)
}
```

## Find some first data
```{r findOne}
if(mongo.is.connected(mongo) == TRUE){
  cityone <- mongo.find.one(mongo, coll, '{"city":"COLORADO CITY"}')
  cityone
  mongo.bson.to.list(cityone)
}
```

## For a long time it was all about creating BSON objects
Since rmongodb version 1.2 you can use JSON directly.
```{r createBSON}
buf <- mongo.bson.buffer.create()
mongo.bson.buffer.append(buf, "city", "COLORADO CITY")
query <- mongo.bson.from.buffer(buf)
query
mongo.bson.from.JSON('{"city":"COLORADO CITY"}')
```

## Find more data
```{r findMore}
if(mongo.is.connected(mongo) == TRUE){
  pop <- mongo.distinct(mongo, coll, "pop")
  hist(pop)
  boxplot(pop)

  mongo.count(mongo, coll, '{"pop":{"$lte":2}}')
  pops <- mongo.find.all(mongo, coll, '{"pop":{"$lte":2}}')
  head(pops)
  dim(pops)
}
```

## Find more data with more compley query
```{r compleyQuery}
library(jsonlite)
json <- '{"pop":{"$lte":2}, "pop":{"$gte":1}}'
cat(prettify(json))
validate(json)
if(mongo.is.connected(mongo) == TRUE){
  mongo.count(mongo, coll, json)
  pops <- mongo.find.all(mongo, coll, json)
  head(pops)
  dim(pops)
}

# still inefficient!
mongo.cursor.to.data.frame
```

## Insert some data to MongoDB
```{r insert}
# insert data
a <- mongo.bson.from.JSON( '{"ident":"a", "name":"Markus", "age":33}' )
b <- mongo.bson.from.JSON( '{"ident":"b", "name":"MongoSoup", "age":1}' )
c <- mongo.bson.from.JSON( '{"ident":"c", "name":"UseR", "age":18}' )

if(mongo.is.connected(mongo) == TRUE){
  icoll <- paste(db, "test", sep=".")
  mongo.insert.batch(mongo, icoll, list(a,b,c) )

  mongo.get.database.collections(mongo, db)
  mongo.find.all(mongo, icoll)
}
```

## Update documents in MongoDB and create indices for efficient queries
```{r update}
if(mongo.is.connected(mongo) == TRUE){
  mongo.update(mongo, icoll, '{"ident":"b"}', '{"$inc":{"age":3}}' )

  mongo.find.all(mongo, icoll)
  
  # Creating an index for the field 'ident'
  mongo.index.create(mongo, icoll, '{"ident":1}')
  # check mongoshell!
}
```

## Drop / Remove collections and databases and Close Connection to MongoDB
```{r dropColls}
if(mongo.is.connected(mongo) == TRUE){
  mongo.drop(mongo, icoll)
  mongo.drop.database(mongo, db)
  mongo.get.database.collections(mongo, db)

  # close connection
  mongo.destroy(mongo)
}
```





# Advanced Steps in rmongodb Package

```{r connect2}
mongo <- mongo.create()
mongo.is.connected(mongo)
```

## Insert Big Data
```{r insertZips2}
if(mongo.is.connected(mongo) == TRUE){
  data(zips)
  head(zips)
  zips[1,]$loc
  
  res <- list(length(dim(zips)[1]))
  for(i in 1:dim(zips)[1]){
    tmp <- zips[i,] 
    res[[i]] <- mongo.bson.from.list(tmp)
  }
  mongo.insert.batch(mongo, "rmongodb.zips", res )
  
  mongo.count(mongo, icoll)
  mongo.find.all(mongo, icoll)
}
```


## MongoDB Aggregation Framework
JSON to BSON not yet working for this secnario
```{r aggregationFramework}
buf <- mongo.bson.buffer.create()
mongo.bson.buffer.start.object(buf, "$group")
mongo.bson.buffer.append(buf, "_id", "$state")
mongo.bson.buffer.start.object(buf, "totalPop")
mongo.bson.buffer.append(buf, "$sum", "$pop")
mongo.bson.buffer.finish.object(buf)
mongo.bson.buffer.finish.object(buf)
bson <- mongo.bson.from.buffer(buf)

bufall <- mongo.bson.buffer.create()
mongo.bson.buffer.append(bufall, "aggregate", "zips")
mongo.bson.buffer.start.array(bufall, "pipeline")
mongo.bson.buffer.append(bufall, "0", bson)
mongo.bson.buffer.finish.object(bufall)
cmd <- mongo.bson.from.buffer(bufall)
cmd
if(mongo.is.connected(mongo) == TRUE){
  res <- mongo.command(mongo, db, cmd)
  res
}
```


## GridFS with rmongodb
GridFS is a specification for storing and retrieving files 
that exceed the BSON-document size limit of 16MB.
```{r gridfs}
if(mongo.is.connected(mongo) == TRUE){
  mgrids <- mongo.gridfs.create(mongo, db, prefix = "fs")
  mongo.gridfs.store.file(mgrids, "faust.txt", "Faust")
  gf <- mongo.gridfs.find(mgrids, "Faust")
  mongo.gridfile.get.length(gf)
  mongo.gridfile.get.chunk.count(gf)
  
  
  
  # close connection
  mongo.drop.database(mongo, db)
  mongo.destroy(mongo)
}
```



