%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction to the rmongodb Package}

Introduction to the rmongodb Package
========================================

MongoDB (www.mongodb.org) is a scalable, high-performance, document-oriented NoSQL database. The **rmongodb** package provides an interface from the statistical software R (www.r-project.org) to MongoDB and back using the mongodb-C library. 

This vignette will provide a first introduction to the **rmongodb** package and a lot of code to get stared. If you need some help to get started with MongoDB please check the resources provided by MongoDB: http://docs.mongodb.org/manual/tutorial/getting-started/


# Installing and Loading rmongodb Package
There is a stable CRAN version of **rmongodb** available:
```{r installRmongodb, eval=FALSE}
install.packages("rmongodb")
```

You can also install the latest development version from the GitHub repository:
```{r installDEV, eval=FALSE}
library(devtools)
install_github("rmongodb", "mongosoup")
```
The installation should be very simple and straightforward. No local MongoDB installation is required. Only if you install from source the RUnit tests will need a local MongoDB installation.

Afterwards you can load the **rmongodb** package as any other package:
```{r loadRmongodb}
library(rmongodb)
```






# Connect R to MongoDB
First of all we have to create a connection to a MongoDB installation. Without parameters we connect to the MongoDB installation running on localhost. Parameters for external installations and user authentication are implemented and documented.
```{r connect2Mongo}
help("mongo.create")
mongo <- mongo.create()
mongo
mongo.is.connected(mongo)
```
It is all the time a good idea to check if there is a working connection to your MongoDB to omit errors.

```{r importZIPdata, echo=FALSE, warning=FALSE, results='hide'}
if(mongo.is.connected(mongo) == TRUE){
  # load some data
  library(jsonlite)
  data(zips)
  res <- list(length(dim(zips)[1]))
  for(i in 1:dim(zips)[1]){
    tmp <- zips[i,] 
    res[[i]] <- mongo.bson.from.list(tmp)
  }
  mongo.insert.batch(mongo, "rmongodb.zips", res )
}
```





# Get Databases and Collections
Get all databases of your MongoDB connection:
```{r getDBs}
if(mongo.is.connected(mongo) == TRUE){
  mongo.get.databases(mongo)
}
```
Get all collection in a specific database of your MongoDB. In this case for the "rmongodb" database.
```{r getColls}
if(mongo.is.connected(mongo) == TRUE){
  db <- "rmongodb"
  mongo.get.database.collections(mongo, db)
}
coll <- "rmongodb.zips"
```
We will use the 'zips' collection in the following examples. The 'zips' collection holds the MongoDB example data set called "Zip Code Data Set" (http://docs.mongodb.org/manual/tutorial/aggregation-zip-code-data-set/). This data set is available as JSON and hold zip code data from the US.





# Get Size of Collection, a Sample Document and Values for a Key
With the command *mongo.count* we can check how many documents are in the collection or in a specific query. More information for all functions can be found in the help files.
```{r count, echo=TRUE}
if(mongo.is.connected(mongo) == TRUE){
  help("mongo.count")
  mongo.count(mongo, coll)
}
```

For running some queries it is important to know some details about the available data. First of all you can run the command *mongo.find.one* to get one document from your collection.
```{r findOneFirst, echo=TRUE}
if(mongo.is.connected(mongo) == TRUE){
  mongo.find.one(mongo, coll)
}
```

The command *mongo.distinct* is going to provide a list of all values for a specific key.
```{r Distinct, echo=TRUE}
if(mongo.is.connected(mongo) == TRUE){
  res <- mongo.distinct(mongo, coll, "city")
  head( res )
}
```






# Find some First Data
Now we can run first queries on our MongoDB. In this case we ask for one document providing zip code data for the city "COLORADO CITY". Please be aware, the output of *mongo.find.one* is a BSON object which can not be used directly for further analyses in R. Using the command *mongo.bson.to.list* an R list object will be created from the BSON object.
```{r findOne}
if(mongo.is.connected(mongo) == TRUE){
  cityone <- mongo.find.one(mongo, coll, '{"city":"COLORADO CITY"}')
  print( cityone )
  mongo.bson.to.list(cityone)
}
```






# Creating BSON objects
For a long time it was all about creating BSON objects in R. This was a very difficult and error-prone task. Since **rmongodb** version 1.2 you can use JSON directly and the syntax of the **rmongodb** package is much more similar to the MongoDB shell.
```{r createBSON}
buf <- mongo.bson.buffer.create()
mongo.bson.buffer.append(buf, "city", "COLORADO CITY")
query <- mongo.bson.from.buffer(buf)
query
```
The same BSON object can be created using one line of code and JSON:
```{r createBSONoneLine}
mongo.bson.from.JSON('{"city":"COLORADO CITY"}')
```





# Find more Data
For real analyses it is important to get more than one document of data from MongoDB. In the first case we use the command *mongo.distict* to get an overview about the population distribution. Than, we check for all cities with less than 2 people living there (errors in the data set?). 
```{r findMore, warning=FALSE}
if(mongo.is.connected(mongo) == TRUE){
  pop <- mongo.distinct(mongo, coll, "pop")
  hist(pop)
  boxplot(pop)

  nr <- mongo.count(mongo, coll, '{"pop":{"$lte":2}}')
  print( nr )
  pops <- mongo.find.all(mongo, coll, '{"pop":{"$lte":2}}')
  print( dim(pops) )
  head(pops)
}
```






# Find more Data with more Complex Query
The analyses gets more interesting creating a more complex query with two arguments. Using the package **jsonlite** you can check and visualize your JSON syntax first. Afterwards we query MongoDB with this JSON query.
```{r compleyQuery}
library(jsonlite)
json <- '{"pop":{"$lte":2}, "pop":{"$gte":1}}'
cat(prettify(json))
validate(json)
if(mongo.is.connected(mongo) == TRUE){
  pops <- mongo.find.all(mongo, coll, json)
  print( dim(pops) )
  head(pops)
}
```
In this case you will get a warning because the construction of the data.frame object is still inefficient and error-prone due to the NoSQL data structure. We are currently working on a solution.
```{r inefficient}
mongo.cursor.to.data.frame
```







# Insert some Data to MongoDB
Another interesting part is inserting data into MongoDB.
```{r insert}
# insert data
a <- mongo.bson.from.JSON( '{"ident":"a", "name":"Markus", "age":33}' )
b <- mongo.bson.from.JSON( '{"ident":"b", "name":"MongoSoup", "age":1}' )
c <- mongo.bson.from.JSON( '{"ident":"c", "name":"UseR", "age":18}' )

if(mongo.is.connected(mongo) == TRUE){
  icoll <- paste(db, "test", sep=".")
  mongo.insert.batch(mongo, icoll, list(a,b,c) )

  dbs <- mongo.get.database.collections(mongo, db)
  print(dbs)
  mongo.find.all(mongo, icoll)
}
```






# Update Documents and create Indices for efficient queries
You can also update your data in MongoDB from R and add indices for more efficient queries.
```{r update}
if(mongo.is.connected(mongo) == TRUE){
  mongo.update(mongo, icoll, '{"ident":"b"}', '{"$inc":{"age":3}}' )

  res <- mongo.find.all(mongo, icoll)
  print(res)
  
  # Creating an index for the field 'ident'
  mongo.index.create(mongo, icoll, '{"ident":1}')
  # check mongoshell!
}
```






# Drop / Remove collections and databases and Close Connection to MongoDB
And there are commands to drop databases and collections in MongoDB. At the end of all your analyses it is a good idea to destroy the connection to your MongoDB.
```{r dropColls}
if(mongo.is.connected(mongo) == TRUE){
  mongo.drop(mongo, icoll)
  mongo.drop.database(mongo, db)
  res <- mongo.get.database.collections(mongo, db)
  print(res)
  
  # close connection
  mongo.destroy(mongo)
}
```


# Feedback and Issues
Please do not hesitate to contact us if there are any issues using **rmongodb**. Issues or pull requests can be submitted via github: https://github.com/mongosoup/rmongodb

